Learning constraints in tabular data

### 2016-05-27 Paper/ECAI/tabular.tex
polished and fixed the conclusions

\subsection{Error detection}
\begin{figure*}[thb]
\begin{subfigure}{.70\textwidth}
  \begin{center}
    \includegraphics[width=\textwidth]{figures/fbi_figure_highlighted.png}
  \end{center}
  \caption{Real world tabular constraint reconstruction: FBI crime statistics}
  \label{fig:fbi}
\end{subfigure}
\hfill
\begin{subfigure}{.29\textwidth}
\begin{center}
    \includegraphics[width=\textwidth]{figures/autocompletion_example.png}
  \end{center}
  \caption{\sergey{Samuel, Tias have a look at it, should be corrected and extended, like with groups indications and formulas?}}
  \label{fig:autocompletion_example}

\end{subfigure}

\end{figure*}

\sergey{Either we need to elaborate on the online detection or remove offline-online and simplify the rest, otherwise it is not clear what is going on}\samuel{Is this still the case?}

@@

The second type of error detection is \textit{offline} and attempts to detect errors in a spreadsheet.
Consider, for example, a use case concerning crime statistics provided by the FBI.
An extract of the spreadsheet is depicted in Figure~\ref{fig:fbi}.
When run on this sheet, our system is able to detect the second sum constraint, however, a missing value prevents the first sum constraint from being learned.
A possible approach to detect such missing or wrong values is to look at a sample of the table that excludes the erroneous row, e.g. trying to leave out each row, and compare the constraints learned on this sample with the constraints learned by adding the row containing the mistake.
The sample needs to large enough such that we have enough confidence that the constraints are correct. For example, in Figure \ref{fig:fbi} all but one row satisfy the constraint, which is a strong indication that there is a constraint violation.
If we define the sample to be the snapshot data and the added row to be the additional input~$\Delta D$, this task can be formulated as online error detection and dealt with in the same way.
%Of course, generating large enough samples and rows (columns) to be added can be done in multiple ways.
%In the limited case of detecting single errors one can leave out every row (column), consider the remaining data to be the sample and
%and the left out row (column) to be new input.

@@

Second, there exist several algorithms in the constraint programming community that induce constraint (programs)
from one or more examples and questions. Two well-known lines of research include ModelSeeker \cite{modelseeker} and Quacq \cite{Quacq}.
The former starts from a single example in the form of a vector of values and then exhaustively looks for the most specific constraints from a large constraint library that hold in the vector. To this aim, it cleverly enumerates different tables from the vector. For instance, if the initial vector was of dimension $1 \times 20$, it would consider rearrangements of size $2 \times 10$, $4 \times 5$, $5 \times 4$, … and Modelseeker would then look for both column, row  and even diagonal constraints. Key differences with our technique are that we assume the tables are given (a reasonable assumption if one starts from a spreadsheet), and also that ModelSeeker’s constraints are the typical ones from the constraint programming community\tias{which is all over integers, we also over strings and real-valued, AND, not one vector but we have many indep. blocks}, which are aimed at combinatorial optimization and scheduling rather than what is typically used in spreadsheets. Quacq \cite{Quacq} and its predecessor Conacq \cite{Conacq} acquire a conjunction of constraints using techniques inspired on Mitchell’s version spaces to process examples, some variants also asking informative questions to the user. As the conjunction for typical constraint programming problems (such as Sudoku) can be quite large, a key problem is to cope with redundant constraints\tias{it looks only at pairwise constraints, so an important factor is how fast the algorithm can converget to the entire constraint set. Redundant and spur also issue for these methods}. As Modelseeker, these systems focus on the types of constraints in combinatorial optimization but unlike ModelSeeker and our approach, they do not assume the variables are organized in tabular form.

Third, our work borrows from the seminal work of Gulwani et al. \cite{flashfill} on program synthesis for tabular data that is incorporated in Microsoft’s Excel. In FlashFill, the end-user might work on a table containing the surnames (column A) and names (column B) of politicians. The first row might contain the values A1 = ``Blair'' and B1 = ``Tony'', and the user might then enter in cell C1 the value ``T.B.''. At that point FlashFill would generate a program that produces the output ``T.B.'' for the inputs ``Blair'' and ``Tony'', would also apply it to automatically generate values for the other politicians (rows) in the table. While it might need an extra example to deal with names such as ``Van Rompuy, Herman'' and determine whether it should output ``H.V.'' or ``H.V.R.'' \tias{only strings?}, FlashFill learns correct programs from very few examples. Flashfill has been extended to e.g. FlashExtract \cite{flashextract} to produce a set of tables starting from a document (a txt, spreadsheet or HTML file) with a set of examples marked for extraction. However, unlike our approach, Flashfill requires the user to identify explicitly the desired inputs/outputs from the function as well as to provide explicitly some positive and sometimes also negative examples.  In contrast, our approach is unsupervised, although it would be interesting to study whether it could be improved through user interaction and through user examples.  Another interesting open issue (for all of these techniques with the exception of logical and relational learning) is how to deal with possibly incomplete or noisy data.

Finally, there are the works on  BayesDB \cite{BayesDB} and Tabular \cite{tabular}, two recent probabilistic modeling languages that have been specifically designed for dealing with relational data in tabular form and that can, as FlashFill, also automatically fill out missing values in a table.  However, unlike the other mentioned approaches, it is based on an underlying probabilistic graphical model that performs probabilistic inference rather than identify “hard” constraints.

@@

\paragraph{Future work}
There are multiple directions for future work on this topic.
Integrating the system in an interactive setting would offer users the possibility to easily receive and provide feedback.
Moreover, the system can be adapted to deal with additional types produce less redundant constraints through the use of (heuristic) filtering or post-processing.

Currently, the system only learns single constraints, however, extending our approach to nested constraints would allow more expressive concepts to be learned.
This shift would bring the approach in line with programming by example.

>>>>

\subsection{Error detection}
\begin{figure*}[thb]
\begin{subfigure}{.70\textwidth}
  \begin{center}
    \includegraphics[width=\textwidth]{figures/fbi_figure_highlighted.png}
  \end{center}
  \caption{Real world tabular constraint reconstruction: FBI crime statistics}
  \label{fig:fbi}
\end{subfigure}
\hfill
\begin{subfigure}{.29\textwidth}
\begin{center}
    \includegraphics[width=\textwidth]{figures/autocompletion_example.png}
  \end{center}
  \caption{\sergey{Samuel, Tias have a look at it, should be corrected and extended, like with groups indications and formulas?}}
  \label{fig:autocompletion_example}
\end{subfigure}
\end{figure*}

@@

The second type of error detection is \textit{offline} and attempts to detect errors in a spreadsheet.
Consider, for example, a use case concerning crime statistics provided by the FBI.
An extract of the spreadsheet is depicted in Figure~\ref{fig:fbi}.
When run on this sheet, our system is able to detect the second sum constraint, however, a missing value prevents the first sum constraint from being learned.
A possible approach to detect such missing or wrong values is to look at a sample of the table that excludes the erroneous row, e.g. trying to leave out each row, and compare the constraints learned on this sample with the constraints learned by adding the row containing the mistake.
The sample needs to be large enough such that we have enough confidence that the constraints are correct. For example, in Figure \ref{fig:fbi} all but one row satisfy the constraint, which is a strong indication that there is a constraint violation.
If we define the sample to be the snapshot data and the added row to be the additional input~$\Delta D$, this task can be formulated as online error detection and dealt with in the same way.
%Of course, generating large enough samples and rows (columns) to be added can be done in multiple ways.
%In the limited case of detecting single errors one can leave out every row (column), consider the remaining data to be the sample and
%and the left out row (column) to be new input.

@@

Second, there exist several algorithms in the constraint programming community that induce constraint (programs)
from one or more examples and questions. Two well-known lines of research include ModelSeeker \cite{modelseeker} and Quacq \cite{Quacq}.
The former starts from a single example in the form of a vector of values and then exhaustively looks for the most specific constraints from a large constraint library that hold in the vector. To this aim, it cleverly enumerates different tables from the vector. For instance, if the initial vector was of dimension $1 \times 20$, it would consider rearrangements of size $2 \times 10$, $4 \times 5$, $5 \times 4$, … and Modelseeker would then look for both column, row  and even diagonal constraints. Key differences with our technique are that we assume the tables are given (a reasonable assumption if one starts from a spreadsheet), and also that ModelSeeker’s constraints are the typical ones from the constraint programming community\tias{which is all over integers, we also over strings and real-valued, AND, not one vector but we have many indep. blocks}, which are aimed at combinatorial optimization and scheduling rather than what is typically used in spreadsheets. Quacq \cite{Quacq} and its predecessor Conacq \cite{Conacq} acquire a conjunction of constraints using techniques inspired on Mitchell's version spaces to process examples, some variants also asking informative questions to the user. As the conjunction for typical constraint programming problems (such as Sudoku) can be quite large, a key problem is to cope with redundant constraints\tias{it looks only at pairwise constraints, so an important factor is how fast the algorithm can converget to the entire constraint set. Redundant and spur also issue for these methods}. As Modelseeker, these systems focus on the types of constraints in combinatorial optimization but unlike ModelSeeker and our approach, they do not assume the variables are organized in tabular form.

Third, our work borrows from the seminal work of Gulwani et al. \cite{flashfill} on program synthesis for tabular data that is incorporated in Microsoft’s Excel. In FlashFill, the end-user might work on a table containing the surnames (column A) and names (column B) of politicians. The first row might contain the values A1 = ``Blair'' and B1 = ``Tony'', and the user might then enter in cell C1 the value ``T.B.''. At that point FlashFill would generate a program that produces the output ``T.B.'' for the inputs ``Blair'' and ``Tony'', would also apply it to automatically generate values for the other politicians (rows) in the table. While it might need an extra example to deal with names such as ``Van Rompuy, Herman'' and determine whether it should output ``H.V.'' or ``H.V.R.'', FlashFill learns correct programs from very few examples. Flashfill has been extended to e.g. FlashExtract \cite{flashextract} to produce a set of tables starting from a document (a txt, spreadsheet or HTML file) with a set of examples marked for extraction. However, unlike our approach, Flashfill requires the user to identify explicitly the desired inputs/outputs from the function as well as to provide explicitly some positive and sometimes also negative examples.  In contrast, our approach is unsupervised, although it would be interesting to study whether it could be improved through user interaction and through user examples.  Another interesting open issue (for all of these techniques with the exception of logical and relational learning) is how to deal with possibly incomplete or noisy data.

Finally, there are the works on  BayesDB \cite{BayesDB} and Tabular \cite{tabular}, two recent probabilistic modeling languages that have been specifically designed for dealing with relational data in tabular form and that can, as FlashFill, also automatically fill out missing values in a table.  However, unlike the other mentioned approaches, it is based on an underlying probabilistic graphical model that performs probabilistic inference rather than identify ``hard'' constraints.

@@

For the future development of the system there are multiple directions. Integrating the system in an interactive setting would offer users the possibility to easily receive and provide feedback.
Moreover, the system can be adapted to deal with additional types and produce less redundant constraints through the use of (heuristic) filtering or post-processing.

Currently, the system only learns single constraints and extending our approach to nested constraints would allow more expressive concepts to be learned. Also, \sname only works on complete rows and columns, allowing partial vectors is an important future direction for developing applications based on the system.
%This shift would bring the approach in line with programming by example.
